---
title: "Measurement of language laterality"
author: "DVM Bishop"
date: "11/07/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(stargazer)
require(tidyverse)
require(yarrr)
library(doBy)
library(psych)
library(knitr)
```

## R Markdown

Data from this project are available on Open Science Framework here: <https://osf.io/pq6wu/>.

These results will be added to manuscript entitled: 'Measurement of language laterality using functional transcranial Doppler ultrasound: does a baseline improve measurement?' authored by Holly Rutherford, Zoe V. J. Woodhead, and Dorothy V. M. Bishop.

This study was pre-registered on: <https://osf.io/6qzex/register/565fb3678c5e4a66b5582f67>.

Preregistered hypotheses:

**Hypothesis 1**: Individuals will be consistent in measures of hemispheric lateralization across different language domains; namely, measures of lateralization (Lateralization Index) will be statistically equivalent across Word Generation and Sentence Generation Tasks.

**Hypothesis 2**: Lateralization indices acquired during the Sentence Generation task will be significantly larger when calculated relative to an active baseline task (number counting, i.e. automatic speech production) than when calculated relative to a resting baseline.


```{r readdata}
#setwd("~/deevybee_repo/holly_rutherford")
alldat <- read.csv('WSLG_AnalysisData.csv')
allsubs <- read.csv('WSLG_Participant_Info.csv')
#str(alldat) #uncomment these lines to check variables in the file
#head(alldat)

#compute percentage of trials marked as omitted before processing
#This number is reported in Results.

totNtrials<-61*31
omittrial <- 100*(totNtrials-sum(alldat[,2:61]))/totNtrials 

```

```{r splithalf}
spear<-c(NA,NA,NA)
 print('Spearman split half reliability')
print("Word Generation")
spear[3]<-cor(alldat$odd_Word,alldat$even_Word,method="spearman")
spear[3]
print("Sentence Generation")
spear[2]<-cor(alldat$odd_Sent,alldat$even_Sent,method="spearman")
spear[2]
print("List Generation")
spear[1]<-cor(alldat$odd_List,alldat$even_List,method="spearman")
spear[1]

spear<-round(spear,2)
```

```{r descriptives}
shortdat <- select(alldat,N_List,N_Sent,N_Word,
                   LI_List,LI_Sent,LI_Word,
                   latency_List,latency_Sent,latency_Word,
                   lat_List,lat_Sent,lat_Word)
stargazer(shortdat[1:9], type = "text") #just for quick viewing

sumtable<-round(describe(shortdat),2)

#create data frame for table 1
table1<-data.frame(matrix(NA,nrow=7,ncol=3))
colnames(table1)<-c('List','Sentence','Word')
rownames(table1)<-c('N trials','LI','Latency','N left','N bilateral','N right','Split-half r')
#Now populate table1

for (i in 1:3){
  table1[1,i]<-paste0(sumtable[i,3]," (",sumtable[i,4],")")
    table1[2,i]<-paste0(sumtable[(3+i),3]," (",sumtable[(3+i),4],")")
     table1[3,i]<-paste0(sumtable[(6+i),3]," (",sumtable[(6+i),4],")")
     table1[4,i]<-length(which(shortdat[,(9+i)]==1))
       table1[5,i]<-length(which(shortdat[,(9+i)]==0))
              table1[6,i]<-length(which(shortdat[,(9+i)]==-1))
              table1[7,i]<-spear[i]
       
}
write.table(table1,paste0('holly_table1.txt'),sep='\t',quote=FALSE)
kable(table1)
#add 3 extra columns for laterality categories: these will be overwritten
shortdat$latcat_Word <-shortdat$lat_Word
shortdat$latcat_Sent <-shortdat$lat_Sent
shortdat$latcat_List <-shortdat$lat_List
for (i in 13:15){
shortdat[,i]<-'L'
w<-which(shortdat[,(i-3)]==0)
shortdat[w,i]<-'B'
w<-which(shortdat[,(i-3)]==-1)
shortdat[w,i]<-'R'
}

shortdat$lat3 <- paste0(shortdat$latcat_Sent,shortdat$latcat_Word,shortdat$latcat_List)
print('3 tasks categorical: Sentence, Word, List')
table(shortdat$lat3)
```


One sample t-tests were used to investigate whether the LIs from each task were significantly different from zero 
```{r ttests}
t3<-t.test(shortdat$LI_Word, y = NULL,
       alternative = c("greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
t2<-t.test(shortdat$LI_Sent, y = NULL,
       alternative = c("greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
t1<-t.test(shortdat$LI_List, y = NULL,
       alternative = c("greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)

t4 <-t.test(shortdat$LI_Sent,shortdat$LI_Word,paired=TRUE,alternative='two.sided')

```

```{r pirates}
#make long form file
nsub<-nrow(alldat)
mylong<-data.frame(matrix(NA,nsub*3,ncol=2))
colnames(mylong)<-c('Task','LI')
mylong$Task<-c(rep('Sentence',31),rep('Word',31),rep('List',31))
mylong$LI[1:nsub]<-alldat$LI_Sent
mylong$LI[(nsub+1):(2*nsub)]<-alldat$LI_Word
mylong$LI[(2*nsub+1):(3*nsub)]<-alldat$LI_List
png(filename="LIpirate.png", width=1500, height=1500,res=300)

pirateplot(LI~Task,data=mylong,point.o=.7,inf.f.o=0,inf.b.o=0)
text(1,6,paste0('t = ', round(t1$statistic,2),'\np = ',round(t1$p.value,3)))
text(2,-3,paste0('t = ', round(t2$statistic,2),'\np < .001'))

text(3,-3,paste0('t = ', round(t3$statistic,2),'\np < .001'))

dev.off()
```
##Difference waveforms for the 3 tasks
Decided we also needed a figure to visualise pattern of activation

```{r plotdiffs}
#fiddle about to create list of subject names
subnames<-seq(1:32)
subnames<-c(subnames[1:24],subnames[26:32])
subnames<-paste0('0',subnames)
subnames[1:9]<-paste0('0',subnames[1:9])
nsub<-length(subnames)
tasklist<-c('List','Sent','Word')
     #set up file for grand avg
     df<-data.frame(matrix(NA,nrow=nrow(plotfile),ncol=1+(nsub*3)))
     colnames(df)[1]<-'time'
     df$time<-plotfile$time
thiscol<-0
for (t in 1:3){ #3 tasks
  #set up plot
  thisfile<-paste0('WSLG_Rawmeans/Rawmeans_001_ListGen.csv')
     plotfile<-read.csv(thisfile)
     plot(plotfile$time,plotfile$meanDiff,type='n',xlab='Time (s)',ylab='L-R blood flow',main=tasklist[t],ylim=c(-6,6))


  for (n in 1:nsub){
    thiscol<-thiscol+1
    thisfile<-paste0('WSLG_Rawmeans/Rawmeans_',subnames[n],'_',tasklist[t],'Gen.csv')
    temp<-read.csv(thisfile)
    lines(temp$time,temp$meanDiff,type='l',col='grey')
    df[,thiscol]<-temp$meanDiff
    colnames(df)[thiscol]<-paste0(tasklist[t],subnames[n])
    
  }
  colrange<-c((thiscol-nsub):thiscol)
  thismean<-rowMeans(df[,colrange])
  lines(temp$time,thismean,type='l',col='blue')
  abline(v=4,lty=2)
  abline(v=14,lty=2)
}


```
```{r intertestcorr}
png(filename="correl3.png", width=2500, height=1000,res=300)
par(mfrow=c(1,3))#plots in one row and 3 columns
mycolours<-c('red','blue')
    plot(alldat$LI_List,alldat$LI_Word,pch=19,xlab='LI List',ylab='LI Word',xlim=c(-6,8),ylim=c(-4,8),col=mycolours[allsubs$Handedness])
    text(2,-3,paste0('r = ',round(cor(alldat$LI_List,alldat$LI_Word,method="spearman"),3)))
     abline(h=0, lty=2)
    abline(v=0, lty=2)
      legend("topleft", inset=.05, title="Handedness",
        c('Left','Right'),col=mycolours,pch=16,cex=.8)
  plot(alldat$LI_List,alldat$LI_Sent,pch=19,xlab='LI List',ylab='LI Sentence',xlim=c(-6,8),ylim=c(-4,8),col=mycolours[allsubs$Handedness])
  abline(h=0, lty=2)
    abline(v=0, lty=2)
      text(2,-3,paste0('r = ',round(cor(alldat$LI_List,alldat$LI_Sent,method="spearman"),3)))
    plot(alldat$LI_Sent,alldat$LI_Word,pch=19,xlab='LI Sentence',ylab='LI Word',xlim=c(-6,8),ylim=c(-4,8),col=mycolours[allsubs$Handedness])
     abline(h=0, lty=2)
    abline(v=0, lty=2)
     text(2,-3,paste0('r = ',round(cor(alldat$LI_Sent,alldat$LI_Word,method="spearman"),3)))
    
     dev.off()
```
```{r blandalt}
png(filename="blandaltman.png", width=1500, height=1250,res=300)
mycolours<-c('red','blue')
plot((shortdat$LI_Word+shortdat$LI_Sent)/2, shortdat$LI_Word-shortdat$LI_Sent, main="Bland-Altman Plot",xlab='Mean LI',ylab='Difference in LI',pch=16,col=mycolours[allsubs$Handedness],cex=.8)
abline(h=2.5,lty=2)
abline(h=-2.5,lty=2)
 legend("topright", inset=.05, title="Handedness",
        c('Left','Right'),col=mycolours,pch=16,cex=.8)
dev.off()

```


```{r makeSentminusList}
alldat$SentListA <-alldat$LI_Sent-alldat$LI_List

#A plot just to explore
plot(alldat$LI_Sent,alldat$SentListA,col=mycolours[allsubs$Handedness],pch=16)

```


```{r othersubtract}
listbase <- read.csv('WSLG_ListBaseline_Data.csv')
alldat$SG_LG_b.wave<-listbase$LI_SGvLG
poimeans <- read.csv('WSLG_POImean_Data.csv')
alldat$SG_LG_c.poi<-poimeans$SentGen-poimeans$ListGen
alldat$SentGen_mean<-poimeans$SentGen

stargazer(alldat[,93:95],type='text')

#One-tailed t-tests to check if LI sig different from zero
t.test(alldat$LI_Sent, y = NULL,
       alternative = c("greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
t.test(alldat$SentListA, y = NULL,
       alternative = c("greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
t.test(alldat$SG_LG_b.wave, y = NULL,
       alternative = c("greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
t.test(alldat$SG_LG_c.poi, y = NULL,
       alternative = c("greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)

#Pre-planned matched t-test to compare if original LI with subtracted LI

t.test(alldat$LI_Sent, alldat$SentListA,
       alternative = c( "less"),
       mu = 0, paired = TRUE, var.equal = FALSE,
       conf.level = 0.975)

```
##Exploratory analyses
1. Is bilateral language a true category, or a consequence of poor reliability?

Language laterality is coded as bilateral when the confidence interval for the LI encompasses zero. This could reflect genuine lack of bias, but it could also occur if the subject had noisy data. We explored this question by comparing the values of the standard error of the LI for those with consistent laterality (L or R) and those classified as bilateral. The standard errors for the lateralised vs nonlateralised were comparable for the SG and LG tasks, but for the WG task, the SE was substantially higher for the bilateral group, suggesting that unreliability played a role in their categorisation as bilateral.

```{r bilat.se}
#compare mean SE for lateralised vs bilateral groups
secols<-c(68,77,86,72,81,90)#ses for the 3 tasks then latcategory for 3 tasks
mylabels<-c('Sentence','List','Word')

for (i in 1:3){
  
  mymodel<-lm(alldat[,secols[i]]~abs(alldat[,secols[(i+3)]]))
  print(mylabels[i])
  print(summary(aov(mymodel)))
  
  beanplot(alldat[,secols[i]]~abs(alldat[,secols[(i+3)]]),main=mylabels[i])
}

```
##2. Is the bimodality of laterality indices an artefact of the method of calculation?




##Session information
```{r sessinfo}
sessionInfo()
```