---
title: "Measurement of language laterality"
author: "DVM Bishop"
date: "11/07/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(stargazer)
require(tidyverse)
require(yarrr)
library(doBy)
library(psych)
library(knitr)
require(BlandAltmanLeh)
```

## R Markdown

Data from this project are available on Open Science Framework here: <https://osf.io/pq6wu/>.
N.B. The scripts in this document read raw data files, so working directory must be set to a location where these can be found.

These results will be added to manuscript entitled: 'Measurement of language laterality using functional transcranial Doppler ultrasound: does a baseline improve measurement?' authored by Holly Rutherford, Zoe V. J. Woodhead, and Dorothy V. M. Bishop.

This study was pre-registered on: <https://osf.io/6qzex/register/565fb3678c5e4a66b5582f67>.

Preregistered hypotheses:

**Hypothesis 1**: Individuals will be consistent in measures of hemispheric lateralization across different language domains; namely, measures of lateralization (Lateralization Index) will be statistically equivalent across Word Generation and Sentence Generation Tasks.

**Hypothesis 2**: Lateralization indices acquired during the Sentence Generation task will be significantly larger when calculated relative to an active baseline task (number counting, i.e. automatic speech production) than when calculated relative to a resting baseline.


```{r readdata}
#setwd("~/deevybee_repo/holly_rutherford")
alldat <- read.csv('WSLG_AnalysisData.csv')
allsubs <- read.csv('WSLG_Participant_Info.csv')
#str(alldat) #uncomment these lines to check variables in the file
#head(alldat)

#compute percentage of trials marked as omitted before processing
#This number is reported in Results.

totNtrials<-61*31
omittrial <- 100*(totNtrials-sum(alldat[,2:61]))/totNtrials 

```

```{r splithalf}
spear<-c(NA,NA,NA)
 print('Spearman split half reliability')
print("Word Generation")
spear[3]<-cor(alldat$odd_Word,alldat$even_Word,method="spearman")
spear[3]
print("Sentence Generation")
spear[2]<-cor(alldat$odd_Sent,alldat$even_Sent,method="spearman")
spear[2]
print("List Generation")
spear[1]<-cor(alldat$odd_List,alldat$even_List,method="spearman")
spear[1]

spear<-round(spear,2)
```

```{r descriptives}
shortdat <- select(alldat,N_List,N_Sent,N_Word,
                   LI_List,LI_Sent,LI_Word,
                   latency_List,latency_Sent,latency_Word,
                   lat_List,lat_Sent,lat_Word)
stargazer(shortdat[1:9], type = "text") #just for quick viewing

sumtable<-round(describe(shortdat),2)

#create data frame for table 1
table1<-data.frame(matrix(NA,nrow=7,ncol=3))
colnames(table1)<-c('List','Sentence','Word')
rownames(table1)<-c('N trials','LI','Latency','N left','N bilateral','N right','Split-half r')
#Now populate table1

for (i in 1:3){
  table1[1,i]<-paste0(sumtable[i,3]," (",sumtable[i,4],")")
    table1[2,i]<-paste0(sumtable[(3+i),3]," (",sumtable[(3+i),4],")")
     table1[3,i]<-paste0(sumtable[(6+i),3]," (",sumtable[(6+i),4],")")
     table1[4,i]<-length(which(shortdat[,(9+i)]==1))
       table1[5,i]<-length(which(shortdat[,(9+i)]==0))
              table1[6,i]<-length(which(shortdat[,(9+i)]==-1))
              table1[7,i]<-spear[i]
       
}
write.table(table1,paste0('holly_table1.txt'),sep='\t',quote=FALSE)
kable(table1)
#add 3 extra columns for laterality categories: these will be overwritten
shortdat$latcat_Word <-shortdat$lat_Word
shortdat$latcat_Sent <-shortdat$lat_Sent
shortdat$latcat_List <-shortdat$lat_List
for (i in 13:15){
shortdat[,i]<-'L'
w<-which(shortdat[,(i-3)]==0)
shortdat[w,i]<-'B'
w<-which(shortdat[,(i-3)]==-1)
shortdat[w,i]<-'R'
}

shortdat$lat3 <- paste0(shortdat$latcat_Sent,shortdat$latcat_Word,shortdat$latcat_List)
print('3 tasks categorical: Sentence, Word, List')
table(shortdat$lat3)
```


One sample t-tests were used to investigate whether the LIs from each task were significantly different from zero. Two-sample t-test to compare LI for Sentence Generation and Word Generation
```{r ttests}
t3<-t.test(shortdat$LI_Word, y = NULL,
       alternative = c("greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
t2<-t.test(shortdat$LI_Sent, y = NULL,
       alternative = c("greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
t1<-t.test(shortdat$LI_List, y = NULL,
       alternative = c("greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)

t4 <-t.test(shortdat$LI_Sent,shortdat$LI_Word,paired=TRUE,alternative='two.sided')

```

```{r pirates}
#make long form file
nsub<-nrow(alldat)
mylong<-data.frame(matrix(NA,nsub*3,ncol=2))
colnames(mylong)<-c('Task','LI')
mylong$Task<-c(rep('Sentence',31),rep('Word',31),rep('List',31))
mylong$LI[1:nsub]<-alldat$LI_Sent
mylong$LI[(nsub+1):(2*nsub)]<-alldat$LI_Word
mylong$LI[(2*nsub+1):(3*nsub)]<-alldat$LI_List
png(filename="LIpirate.png", width=1500, height=1500,res=300)
par(mai = c(.5, .8, .05, 0.05))# mai specifies bottom, left, top and right margins in inches

pirateplot(LI~Task,data=mylong,point.o=.7,inf.f.o=0,inf.b.o=0)
text(1,6,paste0('t = ', round(t1$statistic,2),'\np = ',round(t1$p.value,3)))
text(2,-3,paste0('t = ', round(t2$statistic,2),'\np < .001'))

text(3,-3,paste0('t = ', round(t3$statistic,2),'\np < .001'))

dev.off()
```
##Difference waveforms for the 3 tasks
Figures to visualise pattern of activation. Add colour codes to show if subject categorised as L, B or R.

```{r plotdiffs}
#fiddle about to create list of subject names
subnames<-seq(1:32)
subnames<-c(subnames[1:24],subnames[26:32])
subnames<-paste0('0',subnames)
subnames[1:9]<-paste0('0',subnames[1:9])
nsub<-length(subnames)
tasklist<-c('List','Sent','Word')
titlelist<-c('List','Sentence','Word')
catcol<-vector()
catcol[1]<-which(names(alldat)==c("lat_List"))
catcol[2]<-which(names(alldat)==c("lat_Sent"))
catcol[3]<-which(names(alldat)==c("lat_Word"))

catcolour<-c('lightblue','grey','pink')
     #set up file for grand avg
     df<-data.frame(matrix(NA,nrow=nrow(plotfile),ncol=1+(nsub*3)))
     colnames(df)[1]<-'time'
     df$time<-plotfile$time
thiscol<-0
png(filename="LIdiffwaves.png", width=1000, height=1500,res=300)
par(mfrow=c(3,1),mai = c(.3, .5, 0.5, 0.1))
# mai specifies bottom, left, top and right margins in inches

for (t in 1:3){ #3 task5
  #set up plot
  thisfile<-paste0('WSLG_Rawmeans/Rawmeans_001_ListGen.csv')
     plotfile<-read.csv(thisfile)
     plot(plotfile$time,plotfile$meanDiff,type='n',ylab='L-R blood flow',main=titlelist[t],ylim=c(-6,6),cex.axis=.7)
    mtext("Time (s)", side=1, line=2, cex=.6) #x axis label positioned higher than default (default is line 3 which is not visible given margin changes)
 
catdata<-alldat[,catcol[t]]+2  #convert -1 to 1 to 1:3

  for (n in 1:nsub){
    thiscol<-thiscol+1
    thisfile<-paste0('WSLG_Rawmeans/Rawmeans_',subnames[n],'_',tasklist[t],'Gen.csv')
    temp<-read.csv(thisfile)
   lines(temp$time,temp$meanDiff,col=catcolour[catdata[n]])
    df[,thiscol]<-temp$meanDiff
    colnames(df)[thiscol]<-paste0(tasklist[t],subnames[n])
    
  }
  colrange<-c((thiscol-nsub):thiscol)
  thismean<-rowMeans(df[,colrange])
  lines(temp$time,thismean,type='l',col='blue')
  abline(v=7,lty=2)
  abline(v=17,lty=2)
  abline(h=0,lty=1,col='darkgrey')
}
 legend("bottomleft", inset=.05, title="Laterality category",
        c('Left','Bilateral','Right'),col=catcolour[c(3,2,1)],lty=1,cex=.5)
dev.off()

```
```{r intertestcorr}
png(filename="correl3.png", width=2500, height=1000,res=300)
par(mfrow=c(1,3))#plots in one row and 3 columns
mycolours<-c('red','blue')
    plot(alldat$LI_List,alldat$LI_Word,pch=19,xlab='LI List',ylab='LI Word',xlim=c(-6,8),ylim=c(-4,8),col=mycolours[allsubs$Handedness])
    text(2,-3,paste0('r = ',round(cor(alldat$LI_List,alldat$LI_Word,method="spearman"),3)))
     abline(h=0, lty=2)
    abline(v=0, lty=2)
      legend("topleft", inset=.05, title="Handedness",
        c('Left','Right'),col=mycolours,pch=16,cex=.8)
  plot(alldat$LI_List,alldat$LI_Sent,pch=19,xlab='LI List',ylab='LI Sentence',xlim=c(-6,8),ylim=c(-4,8),col=mycolours[allsubs$Handedness])
  abline(h=0, lty=2)
    abline(v=0, lty=2)
      text(2,-3,paste0('r = ',round(cor(alldat$LI_List,alldat$LI_Sent,method="spearman"),3)))
    plot(alldat$LI_Sent,alldat$LI_Word,pch=19,xlab='LI Sentence',ylab='LI Word',xlim=c(-6,8),ylim=c(-4,8),col=mycolours[allsubs$Handedness])
     abline(h=0, lty=2)
    abline(v=0, lty=2)
     text(2,-3,paste0('r = ',round(cor(alldat$LI_Sent,alldat$LI_Word,method="spearman"),3)))
    
     dev.off()
```
```{r blandalt}
png(filename="blandaltman.png", width=1500, height=1250,res=300)
par(mai = c(.5, .5, .2, 0.2))# mai specifies bottom, left, top and right margins in inches

plot((shortdat$LI_Word+shortdat$LI_Sent)/2, shortdat$LI_Word-shortdat$LI_Sent, main="",xlab='Mean LI',ylab='Difference in LI',pch=16,col=mycolours[allsubs$Handedness],cex=.8)


temp<-shortdat$LI_Word-shortdat$LI_Sent
abline(h=mean(temp),lty=2)
CIhi<-mean(temp)+1.96*sd(temp)
CIlo<-mean(temp)-1.96*sd(temp)
rect(-.5,-2.5,7.5,2.5, col= '#FF003322')
abline(h=CIlo,lty=3)
abline(h=CIhi,lty=3)
 legend("topright", inset=.05, title="Handedness",
        c('Left','Right'),col=mycolours,pch=16,cex=.8)
dev.off()

```


```{r makeSentminusList}
alldat$SentListA <-alldat$LI_Sent-alldat$LI_List

#A plot just to explore
plot(alldat$LI_Sent,alldat$SentListA,col=mycolours[allsubs$Handedness],pch=16)

```


```{r othersubtract}
listbase <- read.csv('WSLG_ListBaseline_Data.csv')
alldat$SG_LG_b.wave<-listbase$LI_SGvLG
poimeans <- read.csv('WSLG_POImean_Data.csv')
alldat$SG_LG_c.poi<-poimeans$SentGen-poimeans$ListGen
alldat$SentGen_mean<-poimeans$SentGen

stargazer(alldat[,93:95],type='text')

#One-tailed t-tests to check if LI sig different from zero
t.test(alldat$LI_Sent, y = NULL,
       alternative = c("greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
t.test(alldat$SentListA, y = NULL,
       alternative = c("greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
t.test(alldat$SG_LG_b.wave, y = NULL,
       alternative = c("greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
t.test(alldat$SG_LG_c.poi, y = NULL,
       alternative = c("greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)

#Pre-planned matched t-test to compare if original LI with subtracted LI

t.test(alldat$LI_Sent, alldat$SentListA,
       alternative = c( "less"),
       mu = 0, paired = TRUE, var.equal = FALSE,
       conf.level = 0.975)

```
##Exploratory analyses
1. Is bilateral language a true category, or a consequence of poor reliability?

Language laterality is coded as bilateral when the confidence interval for the LI encompasses zero. This could reflect genuine lack of bias, but it could also occur if the subject had noisy data. We explored this question by comparing the values of the standard error of the LI for those with consistent laterality (L or R) and those classified as bilateral. The standard errors for the lateralised vs nonlateralised were comparable for the SG and LG tasks, but for the WG task, the SE was substantially higher for the bilateral group, suggesting that unreliability played a role in their categorisation as bilateral.

```{r bilat.se}
#compare mean SE for lateralised vs bilateral groups
secols<-c(68,77,86,72,81,90)#ses for the 3 tasks then latcategory for 3 tasks
mylabels<-c('Sentence','List','Word')

for (i in 1:3){
  
  mymodel<-lm(alldat[,secols[i]]~abs(alldat[,secols[(i+3)]]))
  print(mylabels[i])
  print(summary(aov(mymodel)))
  
  beanplot(alldat[,secols[i]]~abs(alldat[,secols[(i+3)]]),main=mylabels[i])
}

```
##2. Is the bimodality of laterality indices an artefact of the method of calculation?
Look at distribution of scores on ListGen with standard LI and poimeanLI

```{r bimodal.check}
LG1<-alldat$LI_List
LG2<-poimeans$ListGen
png(filename="hist2.png", width=1500, height=800,res=300)
par(mai = c(.5, .5, .5, 0.2))# mai specifies bottom, left, top and right margins in inches

par(mfrow=c(1,2))
par(cex=.5) #was struggling to control size in the plot: this works a treat!
hist(LG1,main='LI from peak',xlab='LI',ylim=c(0,9),breaks=seq(-6,6,by=1),cex=.8,col='green')

hist(LG2,main='LI from mean in POI',xlab='LI',ylim=c(0,9),breaks=seq(-6,6,by=1),cex=.8,col='green')

dev.off()

```
##Visualisation of impact of subtracting List from Sentence

```{r vis.subtract}
png(filename="subtractview.png", width=1500, height=1500,res=300)
par(mai = c(.5, .8, .05, 0.5))# mai specifies bottom, left, top and right margins in inches

#sort by LI on sent
thisorder <- order(alldat$SentListA)

#make a plot frame
for3plot <-select(alldat,LI_List,LI_Sent,SentListA,lat_List)
for3plot$lat_List<-for3plot$lat_List+2 #recode so can use as index for colour
for3plot<-for3plot[thisorder,]
#for3plot$ID <-seq(1:nrow(for3plot))
for3plot<-data.frame(t(for3plot)) #tranpose to easily plot lines
xlabs=c('List \nGeneration','Sentence \nGeneration','Sentence\nminus List')
 plot(for3plot[1:3,1],type='n',ylim=c(-6,10),xaxt='n',ylab='LI',xlab='')
 axis(1, seq(1,3,by=1),xlabs,tck=.01)
 catcolour<-c('blue','darkgrey','red') #for R B and L category on List generation
for (i in 1:ncol(for3plot)){
  
  lines(for3plot[1:3,i],col=catcolour[for3plot[4,i]],lty=2,pch=15,type='o')
}

abline(h=0)
 legend("topleft", inset=.05, title="Laterality on \nList Generation",
        c('Left','Bilateral','Right'),col=catcolour[c(3,2,1)],pch=15,cex=.8,bty='n')
 dev.off()
```
##Session information
```{r sessinfo}
sessionInfo()
```